metric analysis

### 2021-08-18 metric_analysis.Rmd

library(dplyr)

library(readr)
resistance <- read_csv("ED-outputs/results/metric/resistance_values.csv")
resilience <- read_csv("ED-outputs/results/metric/resilience_values.csv")
met <- read_csv("ED-outputs/results/constant_annual_met_data.csv")

@@

```{r, echo=FALSE}
met %>% 
  # hgt is constant; remove
  filter(variable != "hgt") %>%   
  group_by(met, variable) %>% 
  summarise(value = mean(value), .groups = "drop") ->
  met2

library(ggplot2)
theme_set(theme_minimal())
ggplot(met2, aes(met, value)) +
  geom_point() + 
  facet_wrap(~variable, scales = "free_y") +
  theme(axis.text.x = element_text(angle = 90))

library(tidyr)
met2_wide <- pivot_wider(met2, id_cols = "met", names_from = "variable")
pca_met <- prcomp(met2_wide[-1], scale. = TRUE)  # -1 removes met year column 
biplot(pca_met, scale = 0)
pairs(met2_wide[-1])
```

@@

metric_combined %>% 
  pivot_longer(dlwrf:vgrd, values_to = "met_value") ->
  metric_comb_long

@@

```{r, message=FALSE}
# Fit a Random Forest for each variable (across all treatment severities)
# Note the r.f. functions don't play well with tibbles, so change to data frame
library(randomForest)
rc_all <- split(as.data.frame(metric_combined), ~ metric + variable)
rf_all <- lapply(rc_all, function(x) 
  # remove metric, variable, met columns and run RF
  randomForest(value ~ ., data = x[-1:-3], importance = TRUE)
)

@@

```{r}
library(forcats)
f_imp_plot <- function(x) {
  ggplot(x, aes(fct_reorder(met_var, `%IncMSE`), `%IncMSE`)) + 
    geom_col() +
    coord_flip() +
    xlab("Met variable") +
    ggtitle(unique(x$variable))
}
lapply(split(filter(import, metric == "resistance"), ~ variable), f_imp_plot)

@@

```{r}
# Re-do above plots but now for resilience
lapply(split(filter(import, metric == "resilience"), ~ variable), f_imp_plot)
p_pdp <- p_pdp %+% filter(partial, metric == "resilience")
print(p_pdp + ylab("Resilience"))
```

@@

@@

>>>>

library(dplyr)
library(stringr)
library(readr)
resistance <- read_csv("./ED-outputs/results/metric/resistance_values.csv")
resilience <- read_csv("./ED-outputs/results/metric/resilience_values.csv")
met <- read_csv("./ED-outputs/results/constant_annual_met_data.csv")

# The variable table, contains inforamtion about the met variable abreviation with a descrioption. 
var_info <- data.table::data.table(met_var = c('dlwrf', 'nbdsf', 'nddsf', 'vbdsf',
                                    'vddsf', 'prate', 'pres', 'hgt', 
                                    'ugrd', 'vgrd', 'sh', 'tmp', 'GPP', 'grd'), 
                       description = c('downward long wave radiation', 
                                       'near infrared beam downward solar radiation', 
                                       'near IR diffuse downward solar radiation', 
                                       'visible beam downward solar radiation',
                                       'visible diffuse downward solar radiation', 
                                       'precipitation rate', 
                                       'atmospheric pressure', 
                                       'geopotential height', 
                                       'zonal wind',
                                       'meridional wind', 
                                       'specific humidity',
                                       'air temperature', 
                                       'pre-dist. productivity', 'total wind'),  
                       units = c('W m-2', 'W m-2', 'W m-2', 'W m-2',
                                 'W m-2', 'kgH2O m-2 s-1', 'Pa', 'm',
                                 'm s-1', 'm s-1', 'kgH2O kgAir-1', 'K', 'MgC ha-1 year-1', 'm s-1'), stringsAsFactors=FALSE)

# Add 2000 GPP to the met df since it is a "condition". 
ED_output <- read.csv("ED-outputs/results/exp-constant-yr.csv", stringsAsFactors = FALSE) %>% 
  filter(scn == "harvest_0" & variable == "GPP" & year == 2000) %>% 
  select(met, value, variable, value, units = unit)

met %>% 
  filter(variable %in% c("ugrd", "vgrd")) %>% 
  mutate(value = value^2) %>% 
  group_by(year, met) %>% 
  summarise(value = sqrt(sum(value))) %>% 
  ungroup() %>%
  mutate(units = "m s-1", 
         description = "total winnd", 
         variable = "grd") -> 
  wind_df

met %>% 
  filter(!variable %in% c("ugrd", "vgrd")) %>% 
  bind_rows(wind_df, 
            ED_output) ->
  met

@@

```{r, echo=FALSE}
des_order <- c("Air Temperature\n(K)", "Visible Beam Downward Solar Radiation\n(W m-2)",
               "Atmospheric Pressure\n(Pa)", "Visible Diffuse Downward Solar Radiation\n(W m-2)", 
               "Downward Long Wave Radiation\n(W m-2)", "Specific Humidity\n(kgH2O kgAir-1)", 
                "Precipitation Rate\n(kgH2O m-2 s-1)", "Total Wind\n(m s-1)", 
               "Pre-Dist. Productivity\n(MgC ha-1 year-1)")

met %>% 
  # hgt is constant; remove
  filter(variable != "hgt") %>%   
  group_by(met, variable) %>% 
  summarise(value = mean(value), .groups = "drop") ->
  met2

met2 %>% 
  mutate(met_var = variable) %>% 
  inner_join(var_info, by = "met_var") %>%  
  mutate(description = paste0(str_to_title(description), '\n(', units, ')')) %>% 
  filter(description %in% des_order) -> 
  to_plot_met2

met2 %>% 
  filter(met == 2019) -> 
  met2019

met2019 %>% 
  mutate(met_var = variable) %>% 
  left_join(var_info, by = "met_var") %>%  
  mutate(description = paste0(str_to_title(description), '\n(', units, ')')) -> 
  to_plot_2019

to_plot_met2$description <- factor(to_plot_met2$description, des_order, 
                              ordered = TRUE) 
to_plot_2019$description <- factor(to_plot_2019$description, des_order, 
                              ordered = TRUE)

library(ggplot2)
theme_set(theme_minimal())
ggplot(to_plot_met2, aes(met, value)) +
  geom_point() + 
  #geom_point(data = met2019, aes(met, value), color = "#D55E00") +
  facet_wrap(~description, scales = "free_y", labeller = label_wrap_gen(width = 20)) +
  theme(axis.text.x = element_text(angle = 90)) + 
  labs(x = NULL, y = 'Mean annual conditions value') -> 
  plot

ggsave(plot, filename = "./C.analysis/figures/conditions_yr.png", width = 10, height = 6)
plot

library(tidyr)
met2_wide <- pivot_wider(met2, id_cols = "met", names_from = "variable")
pca_met <- prcomp(met2_wide[-1], scale. = TRUE)  # -1 removes met year column 
bit(pca_met, scale = 0)
pairs(met2_wide[-1])
```

@@

metric_combined %>% 
  pivot_longer(dlwrf:vddsf, values_to = "met_value") ->
  metric_comb_long

@@

```{r, message=FALSE}
# Fit a Random Forest for each variable (across all treatment severities)
# Note the r.f. functions don't play well with tibbles, so change to data frame
library(randomForest)
d <- as.data.frame(metric_combined)
rc_all <- split(d, interaction(d$metric, d$variable))
rf_all <- lapply(rc_all, function(x) 
  # remove metric, variable, met columns and run RF
  randomForest(value ~ ., data = x[-1:-3], importance = TRUE)
)

@@

```{r}

library(forcats)
f_imp_plot <- function(x) {
  ggplot(x, aes(fct_reorder(met_var, `%IncMSE`), `%IncMSE`)) + 
    geom_col() +
    scale_x_discrete(labels = function(x) str_wrap(x, width = 25)) +
    coord_flip() +
    xlab(NULL) +
    ggtitle(unique(x$variable)) 
}

import %>%
  left_join(var_info) %>% 
  mutate(met_var = str_to_title(description)) -> 
  import2

d <- filter(import2, metric == "resistance")
xx <- lapply(split(d, d$variable), f_imp_plot)

library(cowplot)
varimp_plot <- plot_grid(xx[["GPP"]], xx[["NEP"]], xx[["NPP"]], xx[["Rh"]])
print(varimp_plot)

ggsave(varimp_plot, filename = "./C.analysis/figures/varimp_plot_resistance.png", width = 10, height = 6)

save_plot("varimp_resistance.png", varimp_plot)

@@

```{r}
# Re-do above plots but now for resilience
d <- filter(import2, metric == "resilience")
xx <- lapply(split(d, d$variable), f_imp_plot)
varimp_plot <- plot_grid(xx[["GPP"]], xx[["NEP"]], xx[["NPP"]], xx[["Rh"]])
print(varimp_plot)
ggsave(varimp_plot, filename = "./C.analysis/figures/varimp_plot_resilience.png", width = 10, height = 6)

save_plot("varimp_resilience.png", varimp_plot)

p_pdp <- p_pdp %+% filter(partial, metric == "resilience")
print(p_pdp + ylab("Resilience"))
```

@@

```{r}

var_info <- data.table::data.table(met_var = c('dlwrf', 'nbdsf', 'nddsf', 'vbdsf',
                                    'vddsf', 'prate', 'pres', 'hgt', 
                                    'ugrd', 'vgrd', 'sh', 'tmp', 'GPP'), 
                       description = c('downward long wave radiation', 
                                       'near infrared beam downward solar radiation', 
                                       'near IR diffuse downward solar radiation', 
                                       'visible beam downward solar radiation',
                                       'visible diffuse downward solar radiation', 
                                       'precipitation rate', 
                                       'atmospheric pressure', 
                                       'geopotential height', 
                                       'zonal wind',
                                       'meridional wind', 
                                       'specific humidity',
                                       'air temperature', 
                                       'pre-dist. productivity'),  
                       units = c('W m-2', 'W m-2', 'W m-2', 'W m-2',
                                 'W m-2', 'kgH2O m-2 s-1', 'Pa', 'm',
                                 'm s-1', 'm s-', 'kgH2O kgAir-1', 'K', 'MgC ha-1 year-1'), stringsAsFactors=FALSE)

var_info$met_var <- factor(var_info$met_var, unique(var_info$met_var), ordered = TRUE)

partial %>% 
  left_join(var_info) %>% 
  mutate(description = paste0(str_to_title(description), '\n(', units, ')')) -> 
  data_partial

# Combined partial dependence plot
data_partial %>% 
  mutate(metric = if_else(metric == "resistance", "Resistance", "Resilience")) ->  
  to_plot

des_order <- c("Air Temperature\n(K)", "Visible Beam Downward Solar Radiation\n(W m-2)",
               "Atmospheric Pressure\n(Pa)", "Visible Diffuse Downward Solar Radiation\n(W m-2)", 
               "Downward Long Wave Radiation\n(W m-2)", "Specific Humidity\n(kgH2O kgAir-1)", 
                "Precipitation Rate\n(kgH2O m-2 s-1)", "Pre-Dist. Productivity\n(MgC ha-1 year-1)")

to_plot$description <- factor(to_plot$description, des_order, 
                              ordered = TRUE)

to_plot %>% 
  ggplot(aes(x, y, color = variable)) + 
  geom_line() + 
  xlab("Conditions variable value") + ylab("Metric value") +
  scale_color_discrete("") +
  #facet_grid(metric ~ description, scales = "free", switch = "y") +
  facet_grid(metric ~ description, 
             labeller = label_wrap_gen(width = 10), scales = "free", switch = "y") +
  theme_bw()+ 
  theme(axis.text.x = element_text(angle = 90)) -> 
  plot

plot

ggsave(plot, filename = "./C.analysis/figures/cond_metric.png", width = 10, height = 6)
```

@@

Interesting that when GPP is included as a condition that could explain the resistance & resilience metrics it really only affect the Rh values.

```{r}
met %>% 
  filter(variable == "GPP") %>% 
  select(met, pre_GPP = value) -> 
  pre_dist_prod

met %>% 
  filter(!variable %in% c("GPP", "hgt")) %>% 
  mutate(description = paste0(str_to_title(description), '\n(', units, ')')) %>% 
  select(met, value, description) -> 
  met_values

equal_breaks <- function(n = 3, s = 0.05, ...){
  function(x){
    # rescaling
    d <- s * diff(range(x)) / (1+2*s)
    signif(seq(min(x)+d, max(x)-d, length=n), digits = 3)
  }
}

met_values %>% 
  left_join(pre_dist_prod) %>% 
  ggplot(aes(value, pre_GPP)) + 
  geom_point() +
  facet_wrap(~description, scales = "free_x", labeller = label_wrap_gen(width = 20)) + 
  labs(y = "Pre-disturbance productivity (MgC ha-1)", 
       x = "Meteorological value") + 
  scale_x_continuous(breaks=equal_breaks(n=3, s=0.05), 
                     expand = c(0.05, 0)) + 
    theme(axis.text.x = element_text(angle=90)) -> 
  plot

ggsave(plot, filename = "./C.analysis/figures/gpp_met.png", width = 10, height = 6)

```
